---
title: "Heritability-02.md"
author: "Garett Slater"
date: "April 21 2022"
output:
  html_document: default
  pdf_document: default
---


```{r knitr setup, include=FALSE,  eval=TRUE, echo=FALSE, warning=FALSE}
library(knitr)
knitr::opts_chunk$set(eval=TRUE, cache=FALSE, message=FALSE, warning=FALSE, 
                      comment = "")
```

```{r data}

# Display the current working directory
getwd();

library(pacman)
p_load(readxl,dplyr,tidyverse,lme4,nadiv,MCMCglmm,QGglmm)


#Upload Trait Files
Sperm<-read_excel("~/OneDrive/Manuscripts/GenesOfDrones/Resources/Expression_Drone_Trait.xlsx",sheet="Sperm")
Weight<-read_excel("~/OneDrive/Manuscripts/GenesOfDrones/Resources/Expression_Drone_Trait.xlsx",sheet="Weight")
Morpho<-read_excel("~/OneDrive/Manuscripts/GenesOfDrones/Resources/Expression_Drone_Trait.xlsx",sheet="Morpho")
ID<-read_excel("~/OneDrive/Manuscripts/GenesOfDrones/Resources/Expression_Drone_Trait.xlsx",sheet="ID")
RFID_out<-read.csv("~/OneDrive/Manuscripts/GenesOfDrones/Resources/RFID_2020Final_out.csv")

ID<-ID%>%dplyr::filter(ID!="Texas" & ID!="Russian"& ID!="Purdue"& ID!="CAC"& ID!=0)
RFID_out<-merge(RFID_out,ID,by="UID")
RFID_out$UTCTime_Final<- with(RFID_out, as.POSIXct(paste(UTCDate, UTCTime), format="%m/%d/%Y %H:%M:%OS"))
var_out<-RFID_out %>% group_by(UID) %>% dplyr::summarise(N_flights=length(flight),mean_flight_timediff=mean(timediff)/60,first_flight_age=min(Drone_Age))
Sperm_1<-merge(Sperm,ID)

```

# Data Description

Drones (\~50/Colony) were raised from 24 mother colonies. These mother colonies were from 5 distinct breeding lines (Line) in the United States. Once the drones emerged, they were marked with RFIDs and placed into a single host colony during July 2021. Drones from all colonies were the same age at emergence so weather did not differ among drones. Drone were collected for sperm and morphometric phenotyping at ages 10,14, and 20. Flights traits were recorded daily so these traits are also dependent on sperm phenotyping age.

## Input Files

There are three input files that correspond to sperm, flight, and morphometric traits

Sperm Input File:

-   Sperm_Conc_2: Sperm concentration (10\^6)

-   Sperm_viability: Viable sperm (%)

<!-- -->

        Sperm_Conc_2 Sperm_viability Age  Colony    Line
    1        7.78228      0.94071146  10 CAC2008      CA
    2        0.68441      0.53932584  10 CAC2002      CA
    3        1.04584      0.49264706  10 CAC2008      CA
    4        1.66104      0.80092593  10 CAC2008      CA
    5        5.39838      0.97008547  10 CAC2002      CA
    6        4.73704      0.95129870  10 CAC2002      CA

Morphometric Input File: (Age is a factor)

            Weight    head_length head_width hindwing_length  hindwing_width  frontwing_length  frontwing_width
        1    0.19       3.640      1.761           7.543          3.204           11.997           3.829
        2    0.21       3.886      1.938           8.278          3.234           12.666           4.033
        3    0.23       3.955      1.765           8.141          3.364           12.489           4.058
        4    0.23       3.963      1.877           8.470          3.438           12.361           3.923
        5    0.22       4.039      1.590           8.269          3.007           12.273           3.827
        6    0.21       3.669      1.952           8.182          3.267           12.233           3.851

Flight Input File: (Age is a factor)

-   N_flights: Averge number of flights per day

-   mean_flight_timediff: Average flight per day

-   first_flight_age: Age of first flight

<!-- -->

             N_flights    mean_flight_timediff  first_flight_age   Colony            Age
    1          17          21.38137255                6             R2005            20
    2          10          17.23166667                8             CAI2004          14
    3          12          55.58194444                7             TX2001           20
    4          15          20.68666667                3             R2005            14
    5           6          40.16666667                8             CAC2007          14
    6           1           0.20000000                8             CAI2004          10

# 1. Generate distribution for each trait

I began by generating distributions for each trait.

## Flight Traits

I looked at distribution of three flight traits:

-   **N_flights**: Average number of flights per day
-   **mean_flight_timediff**: Average flight per day
-   **first_flight_age**: Age of first flight

Flight traits regardless of age.

```{r echo=FALSE}
flight_1<-merge(var_out,ID,by="UID")

par(mfrow = c(1,3))
hist(flight_1$mean_flight_timediff,xlab="Flight Duration(Min)",main=NULL)
hist(flight_1$N_flights,xlab="Number of Flights",main=NULL)
hist(flight_1$first_flight_age,xlab="First Flight Age",main=NULL)
mtext("Histogram of Flight Traits", side = 3, line = -2, outer = TRUE)
```

Flight traits for age 10

```{r echo=FALSE}
flight_1<-merge(var_out,ID,by="UID")
flight_1<-flight_1[which(flight_1$phen_age==10),]

par(mfrow = c(1,3))
hist(flight_1$mean_flight_timediff,xlab="Flight Duration(Min)",main=NULL)
hist(flight_1$N_flights,xlab="Number of Flights",main=NULL)
hist(flight_1$first_flight_age,xlab="First Flight Age",main=NULL)
mtext("Histogram of Flight Traits", side = 3, line = -2, outer = TRUE)
```

Flight traits for age 14

```{r echo=FALSE}
flight_1<-merge(var_out,ID,by="UID")
flight_1<-flight_1[which(flight_1$phen_age==14),]

par(mfrow = c(1,3))
hist(flight_1$mean_flight_timediff,xlab="Flight Duration(Min)",main=NULL)
hist(flight_1$N_flights,xlab="Number of Flights",main=NULL)
hist(flight_1$first_flight_age,xlab="First Flight Age",main=NULL)
mtext("Histogram of Flight Traits", side = 3, line = -2, outer = TRUE)
```

```{r echo=FALSE}
flight_1<-merge(var_out,ID,by="UID")
flight_1<-flight_1[which(flight_1$phen_age==20),]

par(mfrow = c(1,3))
hist(flight_1$mean_flight_timediff,xlab="Flight Duration(Min)",main=NULL)
hist(flight_1$N_flights,xlab="Number of Flights",main=NULL)
hist(flight_1$first_flight_age,xlab="First Flight Age",main=NULL)
mtext("Histogram of Flight Traits", side = 3, line = -2, outer = TRUE)
```

## Sperm Traits

Sperm Traits Regardless of age

```{r echo=FALSE}

par(mfrow = c(1,2))
hist(Sperm_1$Sperm_Conc_2,xlab="Sperm Concentration(10^6)",main=NULL)
hist(Sperm_1$Sperm_viability,xlab="Sperm Viability(%)",main=NULL)
mtext("Histogram of Sperm Traits", side = 3, line = -2, outer = TRUE)
```

Sperm Traits Age 10

```{r echo=FALSE}
Sperm_1<-Sperm[which(Sperm$Age==10),]

par(mfrow = c(1,2))
hist(Sperm_1$Sperm_Conc_2,xlab="Sperm Concentration(10^6)",main=NULL)
hist(Sperm_1$Sperm_viability,xlab="Sperm Viability(%)",main=NULL)
mtext("Histogram of Sperm Traits", side = 3, line = -2, outer = TRUE)
```

Sperm Traits Age 14

```{r echo=FALSE}
Sperm_1<-Sperm[which(Sperm$Age==14),]

par(mfrow = c(1,2))
hist(Sperm_1$Sperm_Conc_2,xlab="Sperm Concentration(10^6)",main=NULL)
hist(Sperm_1$Sperm_viability,xlab="Sperm Viability(%)",main=NULL)
mtext("Histogram of Sperm Traits", side = 3, line = -2, outer = TRUE)
```

Sperm Traits Age 20

```{r echo=FALSE}
Sperm_1<-Sperm[which(Sperm$Age==20),]

par(mfrow = c(1,2))
hist(Sperm_1$Sperm_Conc_2,xlab="Sperm Concentration(10^6)",main=NULL)
hist(Sperm_1$Sperm_viability,xlab="Sperm Viability(%)",main=NULL)
mtext("Histogram of Sperm Traits", side = 3, line = -2, outer = TRUE)
```

## Morphology Traits

```{r echo=FALSE}

Morpho_1<-merge(Morpho,ID)
Morpho_1<-merge(Weight,Morpho_1)

par(mfrow = c(1,3))
hist(Morpho_1$head_length ,xlab="Head Length (mm)",main=NULL)
hist(Morpho_1$Weight,xlab="Weight (mg)",main=NULL)
hist(Morpho_1$frontwing_length,xlab="Frontwing Length (mm)",main=NULL)
mtext("Histogram of Morphology traits", side = 3, line = -2, outer = TRUE)
```

# 2. Generate trait means

## Sperm Traits - By Line

Sperm Traits regardless of age (Mean)

```{r echo=FALSE}
Sperm_1<-merge(Sperm,ID)
Sperm_1 %>%group_by(Line)%>%dplyr::summarise(Sperm_Concentration=mean(Sperm_Conc_2),Sperm_Viability=mean(Sperm_viability))
  
```

Sperm Traits with age (Mean)

```{r echo=FALSE}
Sperm_1<-merge(Sperm,ID)
Sperm_1 %>%group_by(Line,Age)%>%dplyr::summarise(Sperm_Concentration=mean(Sperm_Conc_2),Sperm_Viability=mean(Sperm_viability))
  
```

## Sperm Traits - By Colony

Sperm Traits regardless of age

```{r echo=FALSE}
Sperm_1<-merge(Sperm,ID)
Sperm_1 %>%group_by(Colony)%>%dplyr::summarise(Sperm_Concentration=mean(Sperm_Conc_2),Sperm_Viability=mean(Sperm_viability))
  
```

Sperm Traits with age

```{r echo=FALSE}
Sperm_1<-merge(Sperm,ID)
Sperm_1 %>%group_by(Colony,Age)%>%dplyr::summarise(Sperm_Concentration=mean(Sperm_Conc_2),Sperm_Viability=mean(Sperm_viability))
  
```

## Flight Traits - By Line

Flight Traits regardless of age (Mean)

```{r echo=FALSE}
flight_1 %>%group_by(Line)%>%dplyr::summarise(mean_flight_timediff=mean(mean_flight_timediff),first_flight_age=mean(first_flight_age),N_flights=mean(N_flights))
  
```

```{r echo=FALSE}
flight_1 %>%group_by(Line,phen_age)%>%dplyr::summarise(mean_flight_timediff=mean(mean_flight_timediff),first_flight_age=mean(first_flight_age),N_flights=mean(N_flights))
  
```

## Flight Traits - By Colony

Flight Traits regardless of age (Mean)

```{r echo=FALSE}
flight_1 %>%group_by(Colony)%>%dplyr::summarise(mean_flight_timediff=mean(mean_flight_timediff),first_flight_age=mean(first_flight_age),N_flights=mean(N_flights))
  
```

```{r echo=FALSE}
flight_1 %>%group_by(Colony,phen_age)%>%dplyr::summarise(mean_flight_timediff=mean(mean_flight_timediff),first_flight_age=mean(first_flight_age),N_flights=mean(N_flights))
  
```



## Morphology Traits - Line

```{r echo=FALSE}
Morpho_1 %>% group_by(Line) %>% dplyr::summarise(Weight=mean(Weight),head_length=mean(head_length),frontwing_length=mean(frontwing_length))
  
```


## Morphology Traits - Colony
```{r echo=FALSE}
Morpho_1 %>%group_by(Colony)%>%dplyr::summarise(Weight=mean(Weight),head_length=mean(head_length),frontwing_length=mean(frontwing_length))
  
```

# 3. LMER Variation

For non-repeated traits (measured once per drone's life) - I assume that you have traits measured at different ages so I am leaving in the model the factor effect of age - if this is not true, then remove age - pick one focal trait here so we work out the pipeline and then you can repeat it for other such traits:

## Sperm Traits

### Sperm Concentration
```{r echo=FALSE}
lmer(Sperm_Conc_2 ~ Age +  Line +  (1|Colony), REML = TRUE, data=Sperm_1)
  
```

### Sperm Viability
```{r echo=FALSE}
lmer(Sperm_viability ~ Age +  Line +  (1|Colony), REML = TRUE, data=Sperm_1)
  
```

## Flight Traits

Flight Duration
```{r echo=FALSE}
lmer(mean_flight_timediff ~ phen_age +  Line +  (1|Colony), REML = TRUE, data=flight_1)
  
```


### Number of Flights per day
```{r echo=FALSE}
lmer(N_flights ~ phen_age +  Line +  (1|Colony), REML = TRUE, data=flight_1)
  
```

### First Flight Age
```{r echo=FALSE}
lmer(first_flight_age ~ phen_age +  Line +  (1|Colony), REML = TRUE, data=flight_1)
  
```

# 5. MCM Variation

# Pedigree
```{r echo=FALSE}
#Pedigree for drones
Pedigree<-dplyr::select(ID,"ID","Colony")
 Pedigree$Sire<-NA
 Pedigree$Sex<-1
 Pedigree<-data.frame(Pedigree)
 colnames(Pedigree)<-c("ID","Dam","Sire","Sex")
 Pedigree<-Pedigree[!duplicated(Pedigree$ID), ]
 Pedigree<-Pedigree[which(Pedigree$ID!="Russian"),]
 Pedigree<-Pedigree[which(Pedigree$ID!="Texas"),]
 Pedigree<-Pedigree[which(Pedigree$ID!="Purdue"),]
 Pedigree<-Pedigree[which(Pedigree$ID!="CA"),]
 
 #Mother queen pedigrees
 mother<-data.frame(unique(Pedigree$Dam))
 mother$Dam<-NA
 mother$Sire<-NA
 mother$Sex<-0
 colnames(mother)[1]<-"ID"
 
 Pedigree2<-rbind(mother,Pedigree)
 Pedigree2<- Pedigree2[!duplicated( Pedigree2$ID), ]
 Pedigree3<-Pedigree2[,-4]
 colnames(Pedigree3)<-c("animal","dam","sire")
 
 tmp <- makeS(pedigree = Pedigree2, heterogametic = "1", returnS = TRUE)
 Sinv <- tmp[["Sinv"]]

```
# GLMM


### Sperm Concentration
```{r echo=FALSE}
Sperm_1<-merge(Sperm,ID)

priorP <- list(R = list(V = 1, nu = 1),
G = list(G1 = list(V = 1, nu = 1, alpha.mu = 0, alpha.V = 1000),
G2 = list(V = 1, nu = 1, alpha.mu = 0, alpha.V = 1000)))


Sperm_2<-Sperm_1
colnames(Sperm_2)[1]<-"animal"

Sperm_2$Sperm_Conc_3<-ceiling(Sperm_2$Sperm_Conc_2)
Sperm_2[["Dom"]] <- Sperm_2[["animal"]]


model <- MCMCglmm(Sperm_Conc_3 ~ Age + Line,
random = ~ animal+Dom,
family = "poisson",
prior = priorP,
pedigree =Pedigree3,
ginverse = list(Dom = Sinv), 
data = Sperm_2,
nitt = 100000,
burnin = 10000,
thin = 10)

summary(model)
X <- model[["X"]]
predict <- map(1:nrow(model[["Sol"]]),
~ as.vector(X %*% model[["Sol"]][., ]))

paramsPFE <-
pmap_dfr(list(predict = predict,
var.a = model[["VCV"]][ , "animal"],
var.p = rowSums(model[["VCV"]])),
QGparams,
model = "Poisson.log",
verbose = FALSE)
mean(paramsPFE[["h2.obs"]])
HPDinterval(as.mcmc(paramsPFE[["h2.obs"]]))


```

### Sperm Viability
```{r echo=FALSE}

priorP <- list(R = list(V = 1, nu = 1),
G = list(G1 = list(V = 1, nu = 1, alpha.mu = 0, alpha.V = 1000),
G2 = list(V = 1, nu = 1, alpha.mu = 0, alpha.V = 1000)))


Sperm_2<-Sperm_1
colnames(Sperm_2)[1]<-"animal"
Sperm_2$Sperm_viability<-Sperm_2$Sperm_viability*100
Sperm_2$Sperm_viability<-ceiling(Sperm_2$Sperm_viability)
Sperm_2[["Dom"]] <- Sperm_2[["animal"]]

model <- MCMCglmm(Sperm_viability ~ Age + Line,
random = ~ animal + Dom,
family = "poisson",
prior = priorP,
pedigree = Pedigree3,
ginverse = list(Dom = Sinv), 
data = Sperm_2,
nitt = 100000,
burnin = 10000,
thin = 10)

summary(model)

X <- model[["X"]]
predict <- map(1:nrow(model[["Sol"]]),
~ as.vector(X %*% model[["Sol"]][., ]))

paramsPFE <-
pmap_dfr(list(predict = predict,
var.a = model[["VCV"]][ , "animal"],
var.p = rowSums(model[["VCV"]])),
QGparams,
model = "Poisson.log",
verbose = FALSE)
mean(paramsPFE[["h2.obs"]])
HPDinterval(as.mcmc(paramsPFE[["h2.obs"]]))

  
```

## Flight Traits

Flight Duration
```{r echo=FALSE}
flight_1<-merge(var_out,ID,by="UID")

priorP <- list(R = list(V = 1, nu = 1),
G = list(G1 = list(V = 1, nu = 1, alpha.mu = 0, alpha.V = 1000),
G2 = list(V = 1, nu = 1, alpha.mu = 0, alpha.V = 1000)))

flight_2<-flight_1
colnames(flight_2)[5]<-"animal"
flight_2$mean_flight_timediff<-ceiling(flight_2$mean_flight_timediff)
flight_2[["Dom"]] <- flight_2[["animal"]]

model <- MCMCglmm(mean_flight_timediff ~ phen_age + Line,
random = ~ animal + Dom,
family = "poisson",
prior = priorP,
pedigree =Pedigree3,
ginverse = list(Dom = Sinv), 
data = flight_2,
nitt = 100000,
burnin = 10000,
thin = 10)

summary(model)

X <- model[["X"]]
predict <- map(1:nrow(model[["Sol"]]),
~ as.vector(X %*% model[["Sol"]][., ]))

paramsPFE <-
pmap_dfr(list(predict = predict,
var.a = model[["VCV"]][ , "animal"],
var.p = rowSums(model[["VCV"]])),
QGparams,
model = "Poisson.log",
verbose = FALSE)
mean(paramsPFE[["h2.obs"]])
HPDinterval(as.mcmc(paramsPFE[["h2.obs"]]))
```


### Number of Flights per day
```{r echo=FALSE}

priorP <- list(R = list(V = 1, nu = 1),
G = list(G1 = list(V = 1, nu = 1, alpha.mu = 0, alpha.V = 1000),
G2 = list(V = 1, nu = 1, alpha.mu = 0, alpha.V = 1000)))

flight_2<-flight_1
colnames(flight_2)[5]<-"animal"
flight_2[["Dom"]] <- flight_2[["animal"]]

model <- MCMCglmm(N_flights ~ phen_age + Line,
random = ~ animal + Dom,
family = "poisson",
prior = priorP,
pedigree = Pedigree3,
ginverse = list(Dom = Sinv), 
data = flight_2,
nitt = 100000,
burnin = 10000,
thin = 10)

summary(model)

X <- model[["X"]]
predict <- map(1:nrow(model[["Sol"]]),
~ as.vector(X %*% model[["Sol"]][., ]))

paramsPFE <-
pmap_dfr(list(predict = predict,
var.a = model[["VCV"]][ , "animal"],
var.p = rowSums(model[["VCV"]])),
QGparams,
model = "Poisson.log",
verbose = FALSE)
mean(paramsPFE[["h2.obs"]])
HPDinterval(as.mcmc(paramsPFE[["h2.obs"]]))

```

### First Flight Age
```{r echo=FALSE}

priorP <- list(R = list(V = 1, nu = 1),
G = list(G1 = list(V = 1, nu = 1, alpha.mu = 0, alpha.V = 1000),
G2 = list(V = 1, nu = 1, alpha.mu = 0, alpha.V = 1000)))


flight_2<-flight_1
colnames(flight_2)[5]<-"animal"
flight_2[["Dom"]] <- flight_2[["animal"]]

model <- MCMCglmm(first_flight_age ~ phen_age + Line,
random = ~ animal + Dom,
family = "poisson",
prior = priorP,
pedigree =Pedigree3,
ginverse = list(Dom = Sinv), 
data = flight_2,
nitt = 100000,
burnin = 10000,
thin = 10)

summary(model)

X <- model[["X"]]
predict <- map(1:nrow(model[["Sol"]]),
~ as.vector(X %*% model[["Sol"]][., ]))

paramsPFE <-
pmap_dfr(list(predict = predict,
var.a = model[["VCV"]][ , "animal"],
var.p = rowSums(model[["VCV"]])),
QGparams,
model = "Poisson.log",
verbose = FALSE)
mean(paramsPFE[["h2.obs"]])
HPDinterval(as.mcmc(paramsPFE[["h2.obs"]]))

```

## Morphology

### Weight
```{r echo=FALSE}
  
prior <- list(R = list(V = 1, nu = 0.002),
G = list(G1 = list(V = 1, nu = 0.002),
G2 = list(V = 1, nu = 0.002)))

Morpho_2<-Morpho_1
Morpho_2<-merge(Morpho_2,ID[,c("ID","phen_age","Maturity")])
colnames(Morpho_2)[1]<-"animal"
Morpho_2[["Dom"]] <- Morpho_2[["animal"]]

model <- MCMCglmm(Weight ~ phen_age + Line,
random = ~ animal + Dom,
family = "gaussian",
prior = prior,
pedigree =Pedigree3,
ginverse = list(Dom = Sinv),
data = Morpho_2,
nitt = 100000,
burnin = 10000,
thin = 10)

summary(model)
compute_varpred <- function(beta, design_matrix) {
var(as.vector(design_matrix %*% beta))
}

X <- model[["X"]]
vf <- apply(model[["Sol"]], 1, compute_varpred, design_matrix = X)

vpRE <- rowSums(model[["VCV"]])
vpFE <- rowSums(model[["VCV"]]) + vf
mean(vpRE)
mean(vpFE)

heritFE <- model[["VCV"]][ , "animal"] / vpFE
mean(heritFE)
HPDinterval(heritFE) 

```


### basitarsus length
```{r echo=FALSE}
  
prior <- list(R = list(V = 1, nu = 0.002),
G = list(G1 = list(V = 1, nu = 0.002),
G2 = list(V = 1, nu = 0.002)))

Morpho_2<-Morpho_1
Morpho_2<-merge(Morpho_2,ID[,c("ID","phen_age")])
colnames(Morpho_2)[1]<-"animal"
Morpho_2[["Dom"]] <- Morpho_2[["animal"]]

model <- MCMCglmm(basitarsus_length ~ phen_age + Line,
random = ~ animal + Dom,
family = "gaussian",
prior = prior,
ginverse = list(Dom = Sinv),
pedigree =Pedigree3,
data = Morpho_2,
nitt = 100000,
burnin = 10000,
thin = 10)

summary(model)


compute_varpred <- function(beta, design_matrix) {
var(as.vector(design_matrix %*% beta))
}

X <- model[["X"]]
vf <- apply(model[["Sol"]], 1, compute_varpred, design_matrix = X)

vpRE <- rowSums(model[["VCV"]])
vpFE <- rowSums(model[["VCV"]]) + vf
mean(vpRE)
mean(vpFE)

heritFE <- model[["VCV"]][ , "animal"] / vpFE
mean(heritFE)
HPDinterval(heritFE) 

```


### Frontwing length
```{r echo=FALSE}
  
prior <- list(R = list(V = 1, nu = 0.002),
G = list(G1 = list(V = 1, nu = 0.002),
G2 = list(V = 1, nu = 0.002)))

Morpho_2<-Morpho_1
Morpho_2<-merge(Morpho_2,ID[,c("ID","phen_age")])
colnames(Morpho_2)[1]<-"animal"
Morpho_2[["Dom"]] <- Morpho_2[["animal"]]


model <- MCMCglmm(frontwing_width ~ phen_age + Line,
random = ~ animal + Dom,
family = "gaussian",
prior = prior,
pedigree =Pedigree3,
data = Morpho_2,
ginverse = list(Dom = Sinv),
nitt = 100000,
burnin = 10000,
thin = 10)

summary(model)

compute_varpred <- function(beta, design_matrix) {
var(as.vector(design_matrix %*% beta))
}

X <- model[["X"]]
vf <- apply(model[["Sol"]], 1, compute_varpred, design_matrix = X)

vpRE <- rowSums(model[["VCV"]])
vpFE <- rowSums(model[["VCV"]]) + vf
mean(vpRE)
mean(vpFE)

heritFE <- model[["VCV"]][ , "animal"] / vpFE
mean(heritFE)
HPDinterval(heritFE) 

```


### Maturity
```{r echo=FALSE}
  
priorB <- list(R = list(V = 1, fix = 1),
G = list(G1 = list(V = 1, nu = 1000, alpha.mu = 0, alpha.V = 1),
G2 = list(V = 1, nu = 1000, alpha.mu = 0, alpha.V = 1)))

Morpho_2<-Morpho_1
Morpho_2<-merge(Morpho_2,ID[,c("ID","phen_age","Maturity")])
colnames(Morpho_2)[1]<-"animal"
Morpho_2[["Dom"]] <- Morpho_2[["animal"]]

model <- MCMCglmm(Maturity ~ phen_age + Line,
random = ~ animal + Dom,
family = "threshold",
prior = prior,
pedigree = Pedigree3,
ginverse = list(Dom = Sinv),
data = Morpho_2,
nitt = 100000,
burnin = 10000,
thin = 10)

summary(model)

herit <-model[["VCV"]][ , "animal"] / rowSums(model[["VCV"]])

effectiveSize(herit)
mean(herit)
HPDinterval(herit) 


paramsB <-pmap_dfr(list(mu = model[["Sol"]][ , "(Intercept)"],
var.a = model[["VCV"]][ , "animal"],
var.p = rowSums(model[["VCV"]]) - 1), # Note the - 1 here
QGparams,
model = "binom1.probit",
verbose = FALSE)


mean(paramsB[["h2.obs"]])
HPDinterval(as.mcmc(paramsB[["h2.obs"]]))

```