---
title: "05b_Brain_Module_Trait_Relationship.rmd"
author: "Garett Slater"
date: "April 21 2022"
output:
  html_document: default
  pdf_document: default
---

# Module and Trait Relationship
I looked the relationship between testes gene clusters and important sperm traits, i.e. sperm concentration and sperm viability.


```{r knitr setup, include=FALSE,  eval=TRUE, echo=FALSE, warning=FALSE}
library(knitr)
knitr::opts_chunk$set(eval=TRUE, cache=FALSE, message=FALSE, warning=FALSE, 
                      comment = "")
```


```{r echo=FALSE}
library(pacman)
p_load(readxl,dplyr,tidyverse,lme4,fmsb,WGCNA)

setwd("~/OneDrive/Manuscripts/GenesofDrones/WGCNA")
load(file="Brain-networkConstruction-auto.RData")
load(file="B_datExpr.RData")
load(file="T_datExpr.RData")
Sperm<-read_excel("~/OneDrive/Manuscripts/GenesOfDrones/Resources/Expression_Drone_Trait.xlsx",sheet="Sperm")
ID<-read_excel("~/OneDrive/Manuscripts/GenesOfDrones/Resources/Expression_Drone_Trait.xlsx",sheet="ID")
Weight<-read_excel("~/OneDrive/Manuscripts/GenesOfDrones/Resources/Expression_Drone_Trait.xlsx",sheet="Weight")
Morpho<-read_excel("~/OneDrive/Manuscripts/GenesOfDrones/Resources/Expression_Drone_Trait.xlsx",sheet="Morpho")
RFID_out<-read.csv("~/OneDrive/Manuscripts/GenesOfDrones/Resources/RFID_2020Final_out.csv")

#Flight Traits
RFID_out<-merge(RFID_out,ID,by="UID")
RFID_out$UTCTime_Final<- with(RFID_out, as.POSIXct(paste(UTCDate, UTCTime), format="%m/%d/%Y %H:%M:%OS"))
var_out<-RFID_out %>% group_by(UID) %>% dplyr::summarise(N_flights=length(flight),mean_flight_timediff=mean(timediff)/60,first_flight_age=min(Drone_Age))
```

# Brain Heatmap and Sperm Traits

```{r Sperm}
# Define numbers of genes and samples
nGenes = ncol(B_datExpr);
nSamples = nrow(B_datExpr);
# Recalculate MEs with color labels
MEs0 = moduleEigengenes(B_datExpr, B_moduleColors)$eigengenes
MEs = orderMEs(MEs0)

#Filter Sperm
samp_id<-data.frame(SampleB[,3]);colnames(samp_id)<-"ID2"
Sperm_1<-merge(Sperm,SampleB)
Sperm_1<-dplyr::select(Sperm_1,"ID2","Line","Sperm_Conc_2","Sperm_viability")
Sperm_1<-Sperm_1[!duplicated(Sperm_1$ID2),]

#Filter Modules based on sperm traits
MEs$ID2<-rownames(MEs)
MEs<- filter(MEs, MEs$ID2 %in% Sperm_1$ID2)
MEs<-MEs[!duplicated(MEs$ID2),]
MEs<-MEs[,-20]
sp_MEs<-MEs

moduleTraitCor = cor(MEs, Sperm_1[,c(3,4)], use = "p");
moduleTraitPvalue = corPvalueStudent(moduleTraitCor, nSamples);
write.csv(moduleTraitCor,"moduleTraitCor_B_sperm")
write.csv(moduleTraitPvalue,"moduleTraitPvalue_B_sperm")
```

```{r echo=FALSE}
########################################
##              HeatMap               ## 
########################################
setwd("~/OneDrive/Manuscripts/GenesofDrones/WGCNA")
# Will display correlations and their p-values
textMatrix = paste(signif(moduleTraitCor, 2), "\n(",
                   signif(moduleTraitPvalue, 1), ")", sep = "");
dim(textMatrix) = dim(moduleTraitCor)
par(mar = c(6, 8.5, 3, 3));
# Display the correlation values within a heatmap plot
labeledHeatmap(Matrix = moduleTraitCor,
               xLabels = names(Sperm_1[,c(3,4)]),
               yLabels = names(MEs),
               ySymbols = names(MEs),
               colorLabels = FALSE,
               colors = blueWhiteRed(50),
               textMatrix = textMatrix,
               setStdMargins = FALSE,
               cex.text = 0.5,
               zlim = c(-1,1),
               main = paste("Module-trait relationships"))

pdf(file = "Brain_Trait_Correlation_Sperm.pdf", width = 12, height = 9);
par(mar = c(6, 8.5, 3, 3));
# Display the correlation values within a heatmap plot
labeledHeatmap(Matrix = moduleTraitCor,
               xLabels = names(Sperm_1[,c(3,4)]),
               yLabels = names(MEs),
               ySymbols = names(MEs),
               colorLabels = FALSE,
               colors = blueWhiteRed(50),
               textMatrix = textMatrix,
               setStdMargins = FALSE,
               cex.text = 0.5,
               zlim = c(-1,1),
               main = paste("Module-trait relationships"))
dev.off()
```


# Brain Heatmap and Flight Traits

```{r Flight}
setwd("~/OneDrive/Manuscripts/GenesofDrones/WGCNA")
# Define numbers of genes and samples
nGenes = ncol(B_datExpr);
nSamples = nrow(B_datExpr);
# Recalculate MEs with color labels
MEs0 = moduleEigengenes(B_datExpr, B_moduleColors)$eigengenes
MEs = orderMEs(MEs0)

#Filter Flight
samp_id<-data.frame(SampleB[,3]);colnames(samp_id)<-"ID2"
flight<-merge(var_out,ID)
flight_1<-merge(flight,SampleB)
flight_1<-dplyr::select(flight_1,"ID2","Line","mean_flight_timediff","N_flights","first_flight_age")
flight_1<-flight_1[!duplicated(flight_1$ID2),]

#Filter Modules based on sperm traits
MEs$ID2<-rownames(MEs)
MEs<- filter(MEs, MEs$ID2 %in% flight_1$ID2)
MEs<-MEs[,-20]
fl_MEs<-MEs

moduleTraitCor = cor(MEs, flight_1[,c(3,4,5)], use = "p")
moduleTraitPvalue = corPvalueStudent(moduleTraitCor, nSamples);
write.csv(moduleTraitCor,"moduleTraitCor_B_flight")
write.csv(moduleTraitPvalue,"moduleTraitPvalue_B_flight")
```

```{r echo=FALSE}
########################################
##              HeatMap               ## 
########################################
setwd("~/OneDrive/Manuscripts/GenesofDrones/WGCNA")
# Will display correlations and their p-values
textMatrix = paste(signif(moduleTraitCor, 2), "\n(",
                   signif(moduleTraitPvalue, 1), ")", sep = "");
dim(textMatrix) = dim(moduleTraitCor)
par(mar = c(6, 8.5, 3, 3));
# Display the correlation values within a heatmap plot
labeledHeatmap(Matrix = moduleTraitCor,
               xLabels = names(flight_1[,c(3,4,5)]),
               yLabels = names(MEs),
               ySymbols = names(MEs),
               colorLabels = FALSE,
               colors = blueWhiteRed(50),
               textMatrix = textMatrix,
               setStdMargins = FALSE,
               cex.text = 0.5,
               zlim = c(-1,1),
               main = paste("Module-trait relationships"))

pdf(file = "Brain_Trait_Correlation_FlightTraits.pdf", width = 12, height = 9);
par(mar = c(6, 8.5, 3, 3));
# Display the correlation values within a heatmap plot
labeledHeatmap(Matrix = moduleTraitCor,
               xLabels = names(flight_1[,c(3,4,5)]),
               yLabels = names(MEs),
               ySymbols = names(MEs),
               colorLabels = FALSE,
               colors = blueWhiteRed(50),
               textMatrix = textMatrix,
               setStdMargins = FALSE,
               cex.text = 0.5,
               zlim = c(-1,1),
               main = paste("Module-trait relationships"))
dev.off()

```

# Brain Heatmap and morphology Traits

```{r Morphology}
setwd("~/OneDrive/Manuscripts/GenesofDrones/WGCNA")
# Define numbers of genes and samples
nGenes = ncol(B_datExpr);
nSamples = nrow(B_datExpr);
# Recalculate MEs with color labels
MEs0 = moduleEigengenes(B_datExpr, B_moduleColors)$eigengenes
MEs = orderMEs(MEs0)

#Filter morpho
samp_id<-data.frame(SampleB[,3]);colnames(samp_id)<-"ID2"
morpho<-merge(Morpho,ID)
morpho_1<-merge(morpho,SampleB)
morpho_1<-merge(morpho_1,Weight)
morpho_1<-dplyr::select(morpho_1,"ID2","Line","Maturity","Weight","frontwing_width","head_width")
morpho_1$Maturity<-as.numeric(c("NM" = "0", "M" = "1")[morpho_1$Maturity])
morpho_1<-morpho_1[!duplicated(morpho_1$ID2),]

#Filter Modules based on sperm traits
MEs$ID2<-rownames(MEs)
MEs<- filter(MEs, MEs$ID2 %in% morpho_1$ID2)
MEs<-MEs[,-20]
mor_MEs<-MEs

moduleTraitCor = cor(MEs, morpho_1[,c(3,4,5,6)], use = "p")
moduleTraitPvalue = corPvalueStudent(moduleTraitCor, nSamples);
write.csv(moduleTraitCor,"moduleTraitCor_B_morpho")
write.csv(moduleTraitPvalue,"moduleTraitPvalue_B_morpho")
```

```{r echo=FALSE}
########################################
##              HeatMap               ## 
########################################
setwd("~/OneDrive/Manuscripts/GenesofDrones/WGCNA")
# Will display correlations and their p-values
textMatrix = paste(signif(moduleTraitCor, 2), "\n(",
                   signif(moduleTraitPvalue, 1), ")", sep = "");
dim(textMatrix) = dim(moduleTraitCor)
par(mar = c(6, 8.5, 3, 3));
# Display the correlation values within a heatmap plot
labeledHeatmap(Matrix = moduleTraitCor,
               xLabels = names(morpho_1[,c(3,4,5,6)]),
               yLabels = names(MEs),
               ySymbols = names(MEs),
               colorLabels = FALSE,
               colors = blueWhiteRed(50),
               textMatrix = textMatrix,
               setStdMargins = FALSE,
               cex.text = 0.5,
               zlim = c(-1,1),
               main = paste("Module-trait relationships"))

pdf(file = "Brain_Trait_Correlation_MorphoTraits.pdf", width = 12, height = 9);
par(mar = c(6, 8.5, 3, 3));
# Display the correlation values within a heatmap plot
labeledHeatmap(Matrix = moduleTraitCor,
               xLabels = names(morpho_1[,c(3,4,5,6)]),
               yLabels = names(MEs),
               ySymbols = names(MEs),
               colorLabels = FALSE,
               colors = blueWhiteRed(50),
               textMatrix = textMatrix,
               setStdMargins = FALSE,
               cex.text = 0.5,
               zlim = c(-1,1),
               main = paste("Module-trait relationships"))
dev.off()

```


# Maturity
None of the clusters were significant. I did plot Blue because it was significant for first flight age

```{r echo=FALSE}
########################################
##             N_flights              ## 
########################################

# Define variable weight containing the weight column of datTrait
Maturity = as.data.frame(SampleB$Maturity);
names(Maturity) = "Maturity"
Maturity$Maturity<-as.numeric(c("NM" = "0", "M" = "1")[Maturity$Maturity])
# names (colors) of the modules
modNames = substring(names(MEs), 3)
geneModuleMembership = as.data.frame(cor(B_datExpr, B_MEs, use = "p"));
MMPvalue = as.data.frame(corPvalueStudent(as.matrix(geneModuleMembership), nSamples));

names(geneModuleMembership) = paste("MM", modNames, sep="");
names(MMPvalue) = paste("p.MM", modNames, sep="");
geneTraitSignificance = as.data.frame(cor(B_datExpr, Maturity, use = "p"));
GSPvalue = as.data.frame(corPvalueStudent(as.matrix(geneTraitSignificance), nSamples));
names(geneTraitSignificance) = paste("GS.", names(Maturity), sep="");
names(GSPvalue) = paste("p.GS.", names(Maturity), sep="");


#Light Green Cluster
module = "lightgreen"
column = match(module, modNames);
moduleGenes = B_moduleColors==module;
par(mfrow = c(1,1));
verboseScatterplot(abs(geneModuleMembership[moduleGenes, column]),
                   abs(geneTraitSignificance[moduleGenes, 1]),
                   xlab = paste("Module Membership in", module, "module"),
                   ylab = "Gene significance for body weight",
                   main = paste("Module membership vs. gene significance\n"),
                   cex.main = 1.2, cex.lab = 1.2, cex.axis = 1.2, col = module)

#Turquoise Cluster
module = "turquoise"
column = match(module, modNames);
moduleGenes = B_moduleColors==module;
par(mfrow = c(1,1));
verboseScatterplot(abs(geneModuleMembership[moduleGenes, column]),
                   abs(geneTraitSignificance[moduleGenes, 1]),
                   xlab = paste("Module Membership in", module, "module"),
                   ylab = "Gene significance for body weight",
                   main = paste("Module membership vs. gene significance\n"),
                   cex.main = 1.2, cex.lab = 1.2, cex.axis = 1.2, col = module)
```



# Number of flights
None of the clusters were significant. I did plot Blue because it was significant for first flight age

```{r echo=FALSE}
########################################
##             N_flights              ## 
########################################

# Define variable weight containing the weight column of datTrait
N_flights = as.data.frame(SampleB$N_flights);
names(N_flights) = "N_flights"
# names (colors) of the modules
modNames = substring(names(MEs), 3)
geneModuleMembership = as.data.frame(cor(B_datExpr, B_MEs, use = "p"));
MMPvalue = as.data.frame(corPvalueStudent(as.matrix(geneModuleMembership), nSamples));
names(geneModuleMembership) = paste("MM", modNames, sep="");
names(MMPvalue) = paste("p.MM", modNames, sep="");
geneTraitSignificance = as.data.frame(cor(B_datExpr, N_flights, use = "p"));
GSPvalue = as.data.frame(corPvalueStudent(as.matrix(geneTraitSignificance), nSamples));
names(geneTraitSignificance) = paste("GS.", names(N_flights), sep="");
names(GSPvalue) = paste("p.GS.", names(N_flights), sep="");


#Blue Cluster
module = "blue"
column = match(module, modNames);
moduleGenes = B_moduleColors==module;
par(mfrow = c(1,1));
verboseScatterplot(abs(geneModuleMembership[moduleGenes, column]),
                   abs(geneTraitSignificance[moduleGenes, 1]),
                   xlab = paste("Module Membership in", module, "module"),
                   ylab = "Gene significance for body weight",
                   main = paste("Module membership vs. gene significance\n"),
                   cex.main = 1.2, cex.lab = 1.2, cex.axis = 1.2, col = module)
```


# Flight Duration
None of the clusters were significant. I did plot Blue because it was significant for first flight age
```{r echo=FALSE}

########################################
##            Mean Flights          ## 
########################################

# Define variable weight containing the weight column of datTrait
mean_flight = as.data.frame(SampleB$mean_flight_timediff)
names(mean_flight) = "mean_flight"
# names (colors) of the modules
modNames = substring(names(MEs), 3)
geneModuleMembership = as.data.frame(cor(B_datExpr, B_MEs, use = "p"));
MMPvalue = as.data.frame(corPvalueStudent(as.matrix(geneModuleMembership), nSamples));

names(geneModuleMembership) = paste("MM", modNames, sep="");
names(MMPvalue) = paste("p.MM", modNames, sep="");
geneTraitSignificance = as.data.frame(cor(B_datExpr, mean_flight, use = "p"));
GSPvalue = as.data.frame(corPvalueStudent(as.matrix(geneTraitSignificance), nSamples));
names(geneTraitSignificance) = paste("GS.", names(mean_flight), sep="");
names(GSPvalue) = paste("p.GS.", names(mean_flight), sep="");

#Blue Cluster
module = "blue"
column = match(module, modNames);
moduleGenes = B_moduleColors==module;
par(mfrow = c(1,1));
verboseScatterplot(abs(geneModuleMembership[moduleGenes, column]),
                   abs(geneTraitSignificance[moduleGenes, 1]),
                   xlab = paste("Module Membership in", module, "module"),
                   ylab = "Gene significance for body weight",
                   main = paste("Module membership vs. gene significance\n"),
                   cex.main = 1.2, cex.lab = 1.2, cex.axis = 1.2, col = module)
```


# First Flight Age
None of the clusters were significant. I did plot Blue because it was significant for first flight age
```{r echo=FALSE}
########################################
##            Flight Age         ## 
########################################

# Define variable weight containing the weight column of datTrait
flight_age = as.data.frame(SampleB$first_flight_age)
names(flight_age) = "flight_age"
# names (colors) of the modules
modNames = substring(names(MEs), 3)
geneModuleMembership = as.data.frame(cor(B_datExpr, B_MEs, use = "p"));
MMPvalue = as.data.frame(corPvalueStudent(as.matrix(geneModuleMembership), nSamples));

names(geneModuleMembership) = paste("MM", modNames, sep="");
names(MMPvalue) = paste("p.MM", modNames, sep="");
geneTraitSignificance = as.data.frame(cor(B_datExpr, flight_age, use = "p"));
GSPvalue = as.data.frame(corPvalueStudent(as.matrix(geneTraitSignificance), nSamples));
names(geneTraitSignificance) = paste("GS.", names(flight_age), sep="");
names(GSPvalue) = paste("p.GS.", names(flight_age), sep="");

#Blue Cluster
module = "blue"
column = match(module, modNames);
moduleGenes = B_moduleColors==module;
par(mfrow = c(1,1));
verboseScatterplot(abs(geneModuleMembership[moduleGenes, column]),
                   abs(geneTraitSignificance[moduleGenes, 1]),
                   xlab = paste("Module Membership in", module, "module"),
                   ylab = "Gene significance for body weight",
                   main = paste("Module membership vs. gene significance\n"),
                   cex.main = 1.2, cex.lab = 1.2, cex.axis = 1.2, col = module)
```



